---
title: "Data Analysis - Ames House Prices"
author: "Ian Vetter"
date: "2023-07-20"
output: html_document
---

# The Dataset

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}
library(faraway)
library(dplyr)
library(psych)
library(corrplot)
library(ggplot2)
library(ggcorrplot)
library(lares)
library(reshape2)

housing_data = read.csv("dataset\\AmesHousing.csv")

summary(housing_data)

nrow(housing_data)
ncol(housing_data)

# Using small chunk of dataset for quick testing :)

#housing_data = housing_data[1:100,]

# Coercing categorical predictors into factor variables

housing_data[is.na(housing_data)] = 1

for (i in 1:ncol(housing_data)) {
  if (typeof(housing_data[, i]) == "character") {
    if (length(unique(housing_data[, i])) >= 2) {
      housing_data[, i] = as.factor(housing_data[, i])
    }
    
  }
}

#str(housing_data)
```

First few examples:

```{r}
housing_data$SalePrice[1:10]
housing_data$Lot.Area[1:10]
housing_data$Utilities[1:10]
```


# Collinearity and correlation analysis


```{r}

# Subsetting all the numeric elements of the dataset for collinearity and correlation analysis:

n_idxs = unlist(lapply(housing_data, is.numeric), use.names = FALSE)  
numeric_housing_data = housing_data[, n_idxs]

#str(numeric_housing_data)



# MUST specify use = "complete.obs" argument to ignore NA's in dataset
corrs = round(cor(numeric_housing_data, use="complete.obs"), 2)


# some possible correlation plots?

#corrplot(corrs, method="number")
#ggcorrplot(corrs, lab_size = 0.1)

#corrs

ggplot(melt(corrs), aes(Var1, Var2, fill=value)) +
  geom_tile(height=0.9, width=0.9) +
  scale_fill_gradient2(low="blue", mid="white", high="red") +
  theme_minimal() +
  coord_equal() +
  labs(x="",y="",fill="Corr") +
  theme(axis.text.x=element_text(size=5, angle=45, vjust=1, hjust=1, 
                                 margin=margin(-3,0,0,0)),
        axis.text.y=element_text(size=5, margin=margin(0,-3,0,0)),
        panel.grid.major=element_blank()) 

```

Taking a closer look at some of the most correlated predictors:

```{r}

corr_cross(numeric_housing_data,
  max_pvalue = 0.05,
  top = 15 
)

# A few of these further visualized in pairs:

corr_data = numeric_housing_data[, c("Order", "Yr.Sold", "Garage.Cars", "Garage.Area", "Year.Built", "Garage.Yr.Blt", "Gr.Liv.Area", "TotRms.AbvGrd", "Overall.Qual", "Total.Bsmt.SF", "SalePrice")]

pairs(corr_data, col = "dodgerblue")
```

While some of these high correlation measures are to be expected, such as house year built along garage year built, we can also see some non-trivial patterns start to emerge from the more continuous numeric predictors.


# Basic Regression Models

We'll start out with a few preliminary multiple regression models: A full additive model using all available predictors, a reduced additive model that disregards some of the predictors that exhibit collinearity or low significance, and a backwards AIC-selected model.

```{r, results='hide'}
#sapply(lapply(housing_data, unique), length)


# lm() complains when it runs into "NA" values in dataframe, so just replacing any with 0 for now...


full_additive = lm(SalePrice ~ ., data = housing_data)

additive_lowcorr = lm(SalePrice ~ . - Gr.Liv.Area - Garage.Cars - Total.Bsmt.SF - Garage.Yr.Blt - Order - MS.Zoning - Condition.1 - House.Style - Exterior.1st - Exterior.2nd - Bsmt.Qual - Exter.Cond - Bsmt.Cond - Bsmt.Exposure - BsmtFin.Type.2 - Heating - Electrical - Garage.Type, data = housing_data)

additive_aic = step(additive_lowcorr, direction = "backward")

```

```{r}

anova(additive_lowcorr, full_additive)
anova(additive_aic, full_additive)

```

# Transformations

```{r}
par(mfrow = c(2, 2))    

plot(SalePrice ~ Lot.Area, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Lot area")

plot(SalePrice ~ X1st.Flr.SF, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs First floor area")

plot(SalePrice ~ X2nd.Flr.SF, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Second floor area")

plot(SalePrice ~ Garage.Area, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Garage area")
```

```{r}

par(mfrow = c(2, 2))    

plot(SalePrice ~ Total.Bsmt.SF, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Basement area")

plot(SalePrice ~ Lot.Frontage, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Lot frontage")

plot(SalePrice ~ Overall.Qual, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Overall quality")


plot(SalePrice ~ Overall.Cond, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Overall condition")
```

```{r}

par(mfrow = c(2, 2))   

plot(SalePrice ~ Gr.Liv.Area, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Living area")

plot(SalePrice ~ Full.Bath, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs # of full bathrooms")

plot(SalePrice ~ MS.SubClass, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs MS Subclass")

plot(SalePrice ~ Wood.Deck.SF, data = housing_data, col = "grey", pch = 20, cex = 1.5,
     main = "Sale price vs Deck area")
```

```{r}
#log_p1 = lm(SalePrice ~ . + log(X1st.Flr.SF) + log(Garage.Area) + log(Lot.Frontage), data=housing_data)

log_r1 = lm(log(SalePrice) ~ ., data = housing_data)

#summary(log_p1)
summary(log_r1)
```


# Interactions

```{r}

#int_mod1 = lm(SalePrice ~ (.)^2, data = housing_data)
```




